---
title: "lm_fit for covid infection and temperature"
author: "Claire"
date: "12/17/2021"
output: html_document
---

project aim: I perform a statistical analysis for understanding the effect of the environmental temperature on the exponential growth rate of the cases infected by COVID-19 for Belgium and Luxembourg regions.

Hypothesis: $H_0：\beta_1 =0 ；H1:\beta _1 \neq0$


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(lubridate)
```


# coivd data source and import data
```{r}
weekdata <- read.csv("https://opendata.ecdc.europa.eu/covid19/testing/csv/data.csv", na.strings = "", fileEncoding = "UTF-8-BOM")
```

```{r}
sum(weekdata$new_cases)
weekdata %>% 
 group_by(region_name) %>% 
 summarise(cases_sum = sum(new_cases), cases_max = max(new_cases)) %>% 
 arrange(desc(cases_sum))
```

```{r}
glimpse(weekdata)

```
# covid data for Lux anf Bel

```{r}
Lux <- weekdata[weekdata$region_name == "Luxembourg",]
Blg <-weekdata[weekdata$region_name == "Belgium",]
head(Lux,num=3L)
```


## TEMPRETURE DATA

temperature source for LUX: https://www.ecad.eu/dailydata/index.php
```{r}
LUX_temp <- read_csv("data/TG_SOUID100633.txt")
LUX_temp$DATE <- ymd(LUX_temp$DATE)
LUX_temp$TG <- LUX_temp$TG*0.1
glimpse(LUX_temp)
```


temperature source for Blegium and Greece: https://www.visualcrossing.com/weather-data
```{r}
GR_temp <- read_csv("data/Athens,Greece 2020-01-01 to 2021-12-01.csv")
glimpse(GR_temp)
BLG_temp <- read_csv("data/Belgium 2020-02-01 to 2021-12-01.csv")
glimpse(BLG_temp)
```

# data processing: weekly temperature

```{r}
GR_temp_week <- GR_temp %>%
  group_by(week = ymd(cut(datetime, "week")) )%>% 
  mutate(weekly_temp = sum(temp)/7)%>% 
  mutate(year_week = strftime(datetime, format = "%Y-W%V"))

BLG_temp_week <- BLG_temp %>%
  group_by(week = ymd(cut(datetime, "week")) )%>% 
  mutate(weekly_temp = sum(temp)/7)%>% 
  mutate(year_week = strftime(datetime, format = "%Y-W%V"))
```

# join covid data and temperature data

```{r}

```

# Reading data for Lux and Belgium
```{r}
library(rio)
LUX_original <- import("data/LUX_n.xlsx")
BLG_original <- read_csv("data/BLG.csv")
LUX <- LUX_original %>% 
  filter(new_cases>0)
BLG <- BLG_original %>% 
  filter(new_cases>0)
```
```{r}
glimpse(LUX)
```
```{r}
glimpse(BLG)

```

# Covid cases distribution with time
```{r}
LUX %>% ggplot( aes(date, as.numeric(new_cases))) +
 geom_col(fill = "blue", alpha = 0.6) + 
 theme_minimal(base_size = 14) +
  labs(title = "Weekly COVID-19 Cases in Luxembourg", x= "Date", y= "Daily confirmed cases") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
```
```{r}
BLG %>% ggplot( aes(date, as.numeric(new_cases))) +
 geom_col(fill = "blue", alpha = 0.6) + 
 theme_minimal(base_size = 14) +
  labs(title = "Weekly COVID-19 Cases in Belgium", x= "Date", y= "Daily confirmed cases") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
```

# Weekly confirmed new cases per capita calculation (cpc)

```{r}
lux_population = 626108
blg_population = 11522440

LUX$cpc <- 10000*LUX$new_cases/lux_population
BLG$cpc <- 10000*BLG$new_cases/blg_population
```

# start and end dates for the calculations
```{r}
start_date = as.Date("2020-7-1")
end_date = as.Date("2021-12-15")
```

#  Weekly confirmed new cases per capita with tempreture 

```{r}
LUX %>% 
  filter(date<end_date & date>start_date) %>% 
  filter(weekly_temp>-5) %>% 
  ggplot(aes(date, cpc)) +
  geom_col() +
  geom_point(aes(y = weekly_temp/0.5), color = "red") +
  scale_y_continuous("Daily confirmed new cases per capita", sec.axis = sec_axis( ~.*0.5, name = "Temperature")) +
  ggtitle("The Temperature and Covid afecttion data in Luxembourg") 
```

```{r}
BLG %>% 
  filter(date<end_date & date>start_date) %>% 
  filter(weekly_temp>0) %>% 
ggplot(  aes(date, cpc)) +
  geom_col() +
  geom_point(aes(y = weekly_temp/0.5), color = "red") +
  scale_y_continuous("Weekly confirmed new cases per capita ", sec.axis = sec_axis( ~.*0.5, name = "Temperature")) +
  ggtitle("The Temperature and Covid afecttion data in Belgium") 
```
# Weely confirmed new cases pc vs tempreture

```{r}
LUX$logC <- log(LUX$new_cases)
BLG$logC <- log(BLG$new_cases)

mean(LUX$logC)
mean(BLG$logC)
LUX %>% 
  filter(date<end_date & date>start_date) %>%
  filter(weekly_temp>0) %>% 
  ggplot(aes(weekly_temp, logC))+
  geom_point()
  
```
```{r}
BLG %>% 
  filter(date<end_date & date>start_date) %>%
  filter(weekly_temp>0) %>% # 3 outliers
  ggplot(aes(weekly_temp, logC))+
  geom_point()
  
```

# Linear Model fit
statistic mode
$$ \pmb{Y= \beta_0+\beta_1X+\epsilon \quad where \quad \epsilon \sim N(0,\sigma^2)} $$
$$ Y \sim N(0,\sigma^2)$$
$$\pmb{E}(y_i)=E(\beta_0+\beta_1x_i+\epsilon_i)=\beta_0+\beta_1x_i+E(\epsilon_i)=\beta_0+\beta_1x_i$$

$$Var(y_i)=Var(\beta_0+\beta_1X+\epsilon_i)=Var(\epsilon_i)=\sigma^2$$
## standard model for the evolution of the infected cases by viruses
$$dN/dt=L(N) \cdot F(N),  \quad with  \quad L(N) = \lambda \cdot N, F(N)=1-N^b$$
##where N(t) is the number of total infected cases,Nmax is the maximum possible number of infected cases.

```{r}
LUX_fit <- LUX %>% 
  filter(date<end_date & date>start_date) %>%
  filter(weekly_temp>0) %>% 
  lm(logC ~ weekly_temp, data = .) 

summary(LUX_fit)
confint(LUX_fit)

LUX_C <- LUX %>% 
  filter(date<end_date & date>start_date)
cor(LUX_C$weekly_temp, LUX_C$logC)
cov(LUX_C$weekly_temp, LUX_C$logC)
```

```{r}
BLG_fit <- BLG %>% 
  filter(date<end_date & date>start_date) %>%
  filter(weekly_temp>0) %>% 
  lm(logC ~ weekly_temp, data = .) 

summary(BLG_fit)
confint(BLG_fit)

BLG_C <- BLG %>% 
  filter(date<end_date & date>start_date)
cor(BLG_C$weekly_temp, BLG_C$logC)
cov(BLG_C$weekly_temp, BLG_C$logC)
```
## Fitting model for the data in Luxembourg Ŷ = 7.96 - 0.105∙T
$$\hat{Y}=7.96 - 0.105∙T$$
## Fitting model for the data in Blegium Ŷ = Ŷ = 10.9 - 0.107∙T

$$\hat{Y}=Ŷ = 10.9 - 0.107∙T$$
## estimation analysis
Hypothesis: $$H_0：\beta_1 =0 ；H1:\beta _1 \neq0$$

##Test based on the hyperthesis

```{r}
autoplot(LUX_fit)
```

```{r}
autoplot(BLG_fit)
```
## calculate T and P values. As observed from the above results, calculated P value is less than alpha. This gave us enough evidence with 95% confidence we say that we reject null hypothesis. In other words, with %95 confidence beta value is not zero. So we reject $ H_0：\beta_1 =0 $.

## Fitting model figure plot

```{r}
LUX %>% 
  filter(date<end_date & date>start_date) %>%
  filter(weekly_temp>0 ,new_cases>0) %>% 
  ggplot(aes(weekly_temp, logC))+
  geom_point()+
    geom_abline(slope = coef(LUX_fit)[2], intercept = coef(LUX_fit)[1],linetype=2, color= "blue")+
    geom_hline(yintercept = mean(LUX$logC), color = "darkgreen") +
    geom_vline(xintercept = mean(LUX$weekly_temp), color = "darkgreen")+
  stat_smooth(method = lm,)+
  xlim(0, 40)+
  ggtitle("Luxembourg")
```

```{r}
BLG %>% 
  filter(date<end_date & date>start_date) %>%
  filter(weekly_temp>0 ,new_cases>0) %>% 
  ggplot(aes(weekly_temp, logC))+
  geom_point()+
    geom_abline(slope = coef(BLG_fit)[2], intercept = coef(BLG_fit)[1],linetype=2, color= "blue")+
    geom_hline(yintercept = mean(BLG$logC ), color = "darkgreen") +
    geom_vline(xintercept = mean(BLG$weekly_temp), color = "darkgreen")+
  stat_smooth(method = lm,)+
  xlim(0, 40)+
  ggtitle("Belgium")
```

##  Residual analysis 
Hypothesis: $$H_0: Residual\sim N(0,\sigma^2) ；H1:Residual \quad  !\sim N(0,\sigma^2)$$
```{r}

library(broom)
BLG_temp_case <- BLG %>% 
  filter(date<end_date & date>start_date) %>%
  filter(weekly_temp>0 ,new_cases>0) %>% 
  mutate(expected = augment(BLG_fit)$.fitted) %>% 
  mutate(residual = log(new_cases)-expected) 


ggplot(BLG_temp_case, aes(x=residual)) + 
    geom_histogram(aes(y=..density..),    
                   binwidth=.5,
                   colour="black", fill="light blue") +
    geom_density(alpha=.2, fill="blue")+
    stat_function(
    fun = dnorm,
    args = list(mean = mean(BLG_temp_case$residual), sd = sd(BLG_temp_case$residual)),
    linetype=2,
    size= 1.5,
    colour="red" )+
  labs(
    x = "Residual", y = "Density",
    title = "Residual distribution of fitting data in Blegium"
  )
```

```{r}
library(broom)
LUX_temp_case <- LUX %>% 
  filter(date<end_date & date>start_date) %>%
  filter(weekly_temp>0 ,new_cases>0) %>% 
  mutate(expected = augment(LUX_fit)$.fitted) %>% 
  mutate(residual = log(new_cases)-expected) 

ggplot(LUX_temp_case, aes(x=residual)) + 
    geom_histogram(aes(y=..density..),      
                   binwidth=.5,
                   colour="black", fill="light blue") +
    geom_density(alpha=.2, fill="blue") +
    stat_function(
    fun = dnorm,
    args = list(mean = mean(LUX_temp_case$residual), sd = sd(LUX_temp_case$residual)),
    linetype=2,
    size= 1.5,
    colour="red")+
    labs(
    x = "Residual", y = "Density",
    title = "Residual distribution of fitting data in Luxembourg")

```


# Shapiro Test
```{r}
y.res<-residuals(LUX_fit)
head(y.res)
shapiro.test(y.res)

y2.res<-residuals(BLG_fit)
head(y2.res)
shapiro.test(y2.res)
```
# Test result
# The selected significance level alpha = 0.05 which is less than p value for both cases. This means, p value area coves the alpha completely so we don’t have enough evidence to reject null hypothesis. Hence the residual population is Normal.

#Compare the coeffiencts in Luxembourg ($\hat{\beta}_1$) and Belgium($\hat{\beta}_1$)

```{r}
BLG_sum <- BLG_temp_case %>%
  filter(date<end_date & date>start_date) %>%
  filter(weekly_temp>0 ,new_cases>0) %>% 
  mutate(dif_x = weekly_temp - mean(weekly_temp)) %>%
  mutate(difx2=dif_x^2)%>%
  summarise(Sxx = sum(difx2),
            m2 = mean(residual),
            s2 = sd(residual),
            V_B = sqrt(abs(s2/Sxx)))
head(BLG_sum  )


LUX_sum <- LUX_temp_case %>%
  filter(date<end_date & date>start_date) %>%
  filter(weekly_temp>0 ,new_cases>0) %>% 
  mutate(dif_x = weekly_temp - mean(weekly_temp)) %>%
  mutate(difx2=dif_x^2)%>%
  summarise(Sxx = sum(difx2),
            m2 = mean(residual),
            s2 = sd(residual),
            V_B = sqrt(abs(s2/Sxx)))
head(LUX_sum  )



```
## Visulasition of the coeffiencts in Luxembourg ($\hat{\beta}_1$) and Belgium($\hat{\beta}_1$)
```{r}
set.seed(0109)
dat <-   tibble(
    x = seq(-0.18,-0.035,length.out = 1000),
    y1 = dnorm(x, -0.107, BLG_sum$V_B),
    y2 = dnorm(x, -0.105, LUX_sum$V_B))
head(dat)

  ggplot(dat) +
  geom_area(aes(x,y1),fill = "darkblue", alpha = 0.4)+
  geom_area(aes(x,y2),fill = "red", alpha = 0.4)+
  geom_vline(xintercept = -0.107, color = "blue",lwd=1) +
  geom_vline(xintercept = -0.105, color = "red",lwd=1)+
  labs(title = expression("Comparation the coeffiencts in Luxembourg and Belgium"),
         x = expression(hat(beta)[1]),
         y = "density",
      col = expression(paste('Values of ', mu)))+ 
    annotate("text", x=-0.15, y=10, colour="darkblue", label= expression(paste("Belgium", hat(beta)[1]))) +
    annotate("text", x=-0.08, y=22,colour="darkred", label= expression(paste("Luxembourg", hat(beta)[1])))

```
## Conclusion
# The results clearly support the statistically significant relationship of negative correlation between the average environmental temperature and exponential growth rates of the infected cases.


##Prediction
```{r}
new.dat <- data.frame(weekly_temp=35)
predict(BLG_fit, newdata = new.dat, interval = 'confidence')
```

```{r}

new.dat <- data.frame(weekly_temp=35)
predict(LUX_fit, newdata = new.dat, interval = 'confidence')

```

## Reference
[1] Livadiotis G. Statistical analysis of the impact of environmental temperature on the exponential growth rate of cases infected by COVID-19[J]. PLoS one, 2020, 15(5): e0233875.
[2] Relman D. A. Rapid Expert Consultation on SARS-CoV-2 Survival in Relation to Temperature and Humidity and Potential for Seasonality for the COVID-19 Pandemic. The National Academies Press. 2020; https://doi.org/10.17226/25771



